---
title: "Practice assignment"
output: 
  github_document
---

Welcome to STAT540! 

The point of this assignment is to get you warmed up with git, GitHub, markdown, and R so you're more prepared for the future assignments. The first few seminars will cover all you need to know for completing this practice assignment. 

Note that this practice assignment, just like your 30pt assignment, will need to be done in Rmarkdown (.Rmd) because it requires R code. You can start with the .Rmd version of this file, and fill it in where necessary. Note that Question 1 asks you to add an additional file. 

If you have never used R, some resources can be found via [this page](https://support.rstudio.com/hc/en-us/articles/201141096-Getting-Started-with-R). There are [hundreds of tutorials](https://www.google.ca/search?q=R+tutorial) around the web. 

And if you just want to hack away and try to manage just with the built-in help (type `?` before a command, like `?read.table` and you can even get help for things like `[` or `+` if you surround it with quotes `?"["`), a pro tip is if you get a cryptic error message, just paste it into a web search.

## QUESTION 1: Create a README in your repo (1pt)

Once you have accepted this assignment through the Canvas invite link and have a repository set up (in the format of `STAT540-UBC-2022/intro-assignment-yourgithubID`), you can start personalizing it by creating a README.md file. This would help people who visit your repository understand who you are and what your repository is about. 

Please be sure that your markdown is properly formatted. See this [markdown reference](https://guides.github.com/features/mastering-markdown/) from GitHub if you need help getting started. 

Add some information to this README.md file to introduce yourself to the teaching team. Some things you may wish to include:

- Your name, your program 
- Links to other profiles (Twitter, LinkedIn, personal website, other relevant links, etc)
- What do you hope to achieve in this course
- What this repository is for (example: "This repository contains the intro assignment for STAT540")
- Anything else that you want. You can add emoji, too! 

## QUESTION 2: Data inspection with R (2pt)

When answering the following questions, please always "sandwich" your R code with some text. Ensure fluency by avoiding inserting R code or outputs without any explanation. For example, if the question ask you what x + y given x=1 and y=2, you can answer it like this: 

----

Now we want to find the sum of x and y.

```{r example_R}
x <- 1  # Note how I assign them to variables instead of just writing 1+2
y <- 2 
z <- x + y
z # By doing this, I can print out the output of z. Alternately, you can also do (z <- x+y). 
```

Conclusion: we found that the z is `r z`. <---- (and here I'm using inline R code)

If you are looking at the .Rmd of this file (the Rmarkdown that creates this markdown you're looking at), and click on "Raw", you'll see that I have used inline R code to refer to the variable "z" instead of directly typing "3". This is because if we ever need to change the values of x and y, we could possibly forget to update "3" as well. You can imagine that if you have to change the initial inputs of a long statistical reports, all your numbers in the rest of the report will change. You don't want to have to look through every sentence to make sure that your report is still up-to-date. 

By using variables and inline R code, I can change the value of x and/or y, and rest assure that when I re-knit my Rmarkdown, my conclusion statement will be updated. 

There are some [data sets](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html) in base R for you to play around with. We will explore the Titanic dataset. You can view it by typing `Titanic`. Just like with functions, you can use the question mark `?Titanic` to see more information about the dataset. 

Answer the following questions by subsetting, printing out the result or by graphing. 

### 2.1 Passenger breakdown 

Convert the array into a data frame by using `data.frame()` function. 

- How many children and adults were on Titanic? 
- Were there more female adult or male adult passengers? 

Find the answers these questions by adding code to the code block below.     


#### 2.1.1 Load package and convert data to data frame  

**First, we will load the required package for the functions needed and convert the array into a data frame**

```{r}
suppressMessages(library(tidyverse)) # This package contains functions we will use to answer the questions

titanic <- data.frame(Titanic)
```

The data frame has **`r ncol(titanic)` variables** and **`r nrow(titanic)` observations**.


#### 2.1.2 Determine number of children and adults  

**Next, we will determine the number of children and adults**

```{r}
(child_adult <- titanic %>% # Pipe the data set into the function
  select(Age, Freq) %>% # Select the columns (variables) of interest
  group_by(Age) %>% # Group by Age
  summarize(freq = sum(Freq)))
```

From the analysis, **`r child_adult$freq[1]` children and `r child_adult$freq[2]` adults** were on the Titanic.

----
#### 2.1.3 Determine number of female and male adult passengers  

**Let us now find out if there were more female or male adult passengers**

```{r}
(adult_male_female <- titanic %>% # Pipe the titanic data into the filter function
  filter(Age == "Adult") %>% # Filter the data to only Adult passengers
  group_by(Sex) %>% # Group Adult passengers by Sex
  summarize(n = sum(Freq))) # Summarize the frequency of each group
```
The analysis above shows that there were far **more male adult passengers (`r adult_male_female$n[1]`)** than **female adult passengers (`r adult_male_female$n[2]`)**.


### 2.2 Survival 

Using the same data frame, examine the survival rates.

- Did the children have better survival rate than the adults? 
- Which class of passengers have a better survival rate? (Crew, first class, second class, third class)


Find the answers these questions by adding code to the code block below.  


#### 2.2.1 Compare survival rate of children to adults  

**We will go ahead and find out the survival rates between children and adults to be able to answer the question**

```{r}
(survival_age <- titanic %>% 
    select(Age, Survived, Freq) %>% # To answer the question, select the relevant variables
    group_by(Age, Survived) %>% # Group by Age and Survived
    summarize(survival_by_category = sum(Freq)) %>% # Sum up the number of survivals by Age
    mutate(total_passengers_by_age = sum(survival_by_category)) %>% # Find the total under each Age group
    mutate(survival_rate = survival_by_category/total_passengers_by_age * 100)) # Determine survival rate by age
```
The result shows that **`r survival_age$survival_by_category[2]` (`r round(survival_age$survival_rate[2])`%)** of the children survived whereas **`r survival_age$survival_by_category[4]` (`r round(survival_age$survival_rate[4])`%)** of the adults survived. 
**Therefore, the children had better survival rate than adults.**

#### 2.2.2 Compare survival rate by class

**We will now determine the class of passengers with better survival rate**

```{r}
(survival_class <- titanic %>% 
    select(Class, Survived, Freq) %>% # To answer the question, select the relevant variables
    group_by(Class, Survived) %>% # Group by class and survival
    summarize(survival_by_category = sum(Freq)) %>% # Sum up the number of survivals by class
    mutate(total_passengers_by_class = sum(survival_by_category)) %>% # Find the total under each Age group
    mutate(survival_rate = survival_by_category/total_passengers_by_class * 100)) # Determine survival rate by class
```
1. Out of 325 first class passengers, **`r survival_class$survival_by_category[2]` (`r round(survival_class$survival_rate[2])`%)** survived 
2. Out of 285 second class passengers, **`r survival_class$survival_by_category[4]` (`r round(survival_class$survival_rate[4])`%)** survived 
3. Out of 706 third class passengers, **`r survival_class$survival_by_category[6]` (`r round(survival_class$survival_rate[6])`%)** survived
4. Out of 885 crew members, **`r survival_class$survival_by_category[8]` (`r round(survival_class$survival_rate[8])`%)** survived.

**Therefore, first class passengers had the highest survival rate.**


## 3. Data visualization (1.5pt)

Here you'll practice reading data from a text file and do some graphing. The file ["guinea_pig_tooth_growth.txt"](https://raw.githubusercontent.com/STAT540-UBC/STAT540-UBC.github.io/master/homework/intro_assignment/guinea_pigs_tooth_growth.txt) is actually the same data as the dataset `ToothGrowth`. You can use `?ToothGrowth` to see what it's about. However, use `read.table` to read this file into a data frame for this question. 

- Create a figure for this dataset (choose whatever graph you'd like!)
- Explain how your graph is informative: what does it tell you about the result of the experiment? Also explain why you choose to present the data in this way. 

Add code to the code block below to complete these tasks. 

#### 3.1 Read data

**First, we will read the data into R as a data frame**
```{r}
pig_tooth_growth <- data.frame(read.table("guinea_pigs_tooth_growth.txt", header = TRUE)) # Assign the data to a variable "pig_tooth_growth"
```


#### 3.2 Explore data  

**Next, we will explore the data to understand more about it**

```{r}
str(pig_tooth_growth) # Provides information about the data
```

We can see that there are *`r ncol(pig_tooth_growth)` variables* and *`r nrow(pig_tooth_growth)` observations* in the data. The `len` and `dose` variables are `numeric` while the `supp` variable is a `character`.


#### 3.3 Determine effect of dosage and supplement type on tooth growth

**Let us see if increase in dosage and supplement type (VC or OJ) resulted in higher length of odontoblasts(cells responsible for tooth growth)** 

*So, the length of odontoblasts translates to  tooth growth*

We will use a scatterplot to visualize the data

```{r}
pig_tooth_growth %>% # Pipe the data into the ggplot function
    ggplot(aes(dose, len, color = supp)) + # Map the variables to aesthetics
    geom_jitter(width = 0.1) + # A little jitter to avoid over-plotting
    theme_bw() + # Reduce the background
    labs(title = "Effect of Vit. C Dosage and Source on Tooth Growth",
          x = "Dose of Vitamin C",
          y = "Length of odontoblasts (tooth growth)",
          color = "Source of Vit. C") # Give informative labels to the plot
```

This graph is very informative. It shows that there was increase in tooth growth as the dosage of vitamin C increased. It also shows that especially at doses 0.5 and 1.0, OJ(orange juice produced more tooth growth) than VC(ascorbic acid). I wonder why there was no dose at 1.5. The graph clearly shows that. I chose to present the data in this way because it gives me information on individual pigs.


## 4. Mechanics (0.5pt)

This check is for the "quality" of this assignment.  Check out the [assignment tips](https://stat540-ubc.github.io/submission_guide.html#general-tips-for-working-on-the-assignments) if you're not sure what that means. Basically, it's important to write up organized reports that are clear and easy to follow, so that your collaborators and the future you can understand what you are doing. The elements that make up a good report include (not limited to): 

- Properly labeled graphs
- Well-formatted markdown (use of headings, subheadings, and numbering that makes sense)
- No inexplicable R outputs (example: printing out an entire dataframe unnecessarily when one can simply use `str()`) 
- Codes are commented when needed
- No broken code. 

We will also evaluate this part by looking at :

- Rmarkdown is properly knitted: markdown is created and easy to read on GitHub, figure folder is there, etc. 
- all the files associated with this assignment are present in a subfolder (this happens automatically by default). 

Be sure to commit and push your completed assignment before the deadline. Anything submitted or modified after the deadline will not be marked.